---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r}
setwd("F:\\temp\\repdata-data-activity");
activitydat <- read.csv("activity.csv", stringsAsFactors=FALSE);
print(head(activitydat));
```


## What is mean total number of steps taken per day?
```{r results="asis"}
activitydaytotal <- aggregate(activitydat[,1], list(activitydat$date), sum, na.rm=TRUE);
names(activitydaytotal) =  c("Date","Steps");
barplot(activitydaytotal$Steps, names.arg=activitydaytotal$Date, plot=TRUE);
library(xtable);
summary <- data.frame(c("Mean", "Median"), c(mean(activitydaytotal$Steps, na.rm=TRUE), median(activitydaytotal$Steps, na.rm=TRUE)));
names(summary) <- c("    Stat    ", "    Value    ");
xt<-xtable(summary);
print(xt, type="html")
```

## What is the average daily activity pattern?
```{r results="asis"}
activityintervalmean <- aggregate(activitydat[,1], list(activitydat$interval), mean, na.rm=TRUE);
names(activityintervalmean) =  c("Interval","Steps");
plot(activityintervalmean$Interval, activityintervalmean$Steps, type = "l", xlab="Interval", ylab= "Steps(Average)");
axis(side=1, cex.axis=0.35)
```

### 5 minute Interval with the highest number of steps
```{r results="asis"}
activityintervalmean$Interval[activityintervalmean$Steps==max(activityintervalmean$Steps)]
```

## Inputing missing values

### Number of records with missing values is 
```{r results="asis"}
sum(is.na(activitydat$steps));

```
All NA values are replaced using the average for that time interval
  
### Replace all the NA values with intereval mean. Count NA values again
```{r results="asis"}
activitydaymean <- aggregate(activitydat[,1], list(activitydat$date), mean, na.rm=TRUE);
activitydaymean$x[is.nan(activitydaymean$x)] <- 0;
library(dclone);
activitydat1 <- dclone(activitydat);
sum(is.na(activitydat1$steps));
for(i in which(is.na(activitydat[, 1]))){ activitydat1[i,1] = activityintervalmean$Steps[which(activityintervalmean$Interval==activitydat[i, 3])] };
print("New count of NA values in data after replacement of NA's with mean")
sum(is.na(activitydat1$steps));
```

### Plot Histogram, Calculate mean, median again after replacing NA's with interval mean
```{r results="asis"}
activitydaytotal1 <- aggregate(activitydat1[,1], list(activitydat1$date), sum, na.rm=TRUE);
names(activitydaytotal1) =  c("Date","Steps");
barplot(activitydaytotal1$Steps, names.arg=activitydaytotal1$Date, plot=TRUE);
library(xtable);
summary1 <- data.frame(c("Mean", "Median"), c(mean(activitydaytotal1$Steps, na.rm=TRUE), median(activitydaytotal1$Steps, na.rm=TRUE)));
names(summary1) <- c("     Stat1    ", "    Value1    ");
xt1<-xtable(summary1);
print(xt1, type="html")
```

## Are there differences in activity patterns between weekdays and weekends?
```{r results="asis"}
activitydat["weekday"] <- sapply(activitydat$date, function(x){ day <- weekdays(as.Date(x)); if(day == "Saturday" || day == "Sunday")  "WEEKEND" else "WEEKDAY"});
activitydatWD <- subset(activitydat, activitydat$weekday == "WEEKDAY");
activityintervalmeanWD <- aggregate(activitydatWD[,1], list(activitydatWD$interval), mean, na.rm=TRUE);
names(activityintervalmeanWD) =  c("Interval","Steps");
print("Week day graph");
plot(activityintervalmeanWD$Interval, activityintervalmeanWD$Steps, type = "l", xlab="Interval", ylab= "Steps(Average)");
axis(side=1, cex.axis=0.35)

activitydatWE <- subset(activitydat, activitydat$weekday == "WEEKEND");
activityintervalmeanWE <- aggregate(activitydatWE[,1], list(activitydatWE$interval), mean, na.rm=TRUE);
names(activityintervalmeanWE) =  c("Interval","Steps");
par(mfrow=c(1,2)) 

print("Week day graph");
plot(activityintervalmeanWD$Interval, activityintervalmeanWD$Steps, type = "l", xlab="Interval", ylab= "Steps(Average)", main="Weekday graph");

print("Week end graph");
plot(activityintervalmeanWE$Interval, activityintervalmeanWE$Steps, type = "l", xlab="Interval", ylab= "Steps(Average)", main="Weekend graph");

axis(side=1, cex.axis=0.35)

```

#### Week end graph looks different from week day graph